The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-01-12 paper/ms.tex
Last Version, some typos/labels/captions corrected

A quantitative description of the \ly line has been reached through
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become popular due to the
availability of computing power. Early into the 21st century the first
studies focused on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}. Later on the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06}. Similar codes have applied
these results to semi-analytic models of galaxy formation \citep{Orsi12} and
results of large hydrodynamic simulations
\citep{CLARA,Forero12,Behrens13}. Recently Monte Carlo codes have also
been applied to the results of high resolution hydrodynamic
simulations of individual
galaxies\citep{Laursen09,Barnes11,Verhamme12,Yajima12}. While, recent
developments have been focused on the systematic study of clumpy
outflows \citep{DijkstraKramer}and anisotropic velocity configurations
\citep{Zheng2013}.

@@

Performing that study is the main goal of this paper. We investigate for the
first time the impact of rotation on the morphology of the \ly line. We focus on a simplified system, a spherical gas cloud
with homogeneous density and solid body rotation, to study the line
morphology and the escape  fraction in the presence of dust. We base
our work on two independent Monte Carlo based radiative transfer codes
CLARA \citep{CLARA} and XX \citep{DijkstraKramer} .

This paper is structured as follows. In \S
\ref{sec:implementation} we present the implementation of bulk
rotation into the Monte Carlo codes, paying special attention to coordinate
definitions. We also list the different physical parameters in
the simulated grid of models. In the next \S \ref{sec:results} we present the
results of the simulations, with special detail to quantities in the
line that show a clear evolution as a function of the sphere
rotational velocity. In \S \ref{sec:discussion} we discuss the
implications of our results in the interpretation of LAEs
observations and further refinements that can be implemented in the
theoretical modeling of the \ly line. In the last section we present
our conclusions.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NewWidthVvsVmax.png}
    \caption{Half-width for the non-dusty models as a function of
      rotational velocity $V_{\rm max}$. Continuous (dashed) lines
      correspond to homogeneous (central) source
      distributions. \label{fig:widthvsvelocity}} 
\end{figure}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumVvsVmax.png}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  homogeneous (central) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Logarithm of the average number of scatterings as function of
  rotational velocity velocity. Continuous (dashed) lines represent an
  homogeneous (central) distribution of sources. \label{fig:Nscatt}}    
\end{figure}

@@

\begin{figure}
  \includegraphics[width=0.45\textwidth]{escapefraction.png}
   \caption{Escape fraction as a function of rotational velocity, for these models we took $\tau_{a}=1$. The
     continuous (dashed) lines correspond to homogeneous (central)
     models.
     \label{fig:efvsv}}
\end{figure}

@@

\begin{figure}
  \includegraphics[scale=0.35]{muhisto.png}
  \caption{$\mu$ histogram showing the orientation effects for the different
  distributions Central(Points line), Homogeneous(Dashed) and Off-Center(Solid line).  
    \label{fig:muhisto}} 
\end{figure}

@@

\begin{figure*}
\includegraphics[scale=0.3]{OC_profiles.png}
\caption{\ly profiles for the off-center positions. Dashed/solid
  lines represents spheres centered at $-0,5R/+0.5R$.
   \label{fig:OC_profiles}} 
\end{figure*}

@@

\begin{figure*}
\includegraphics[scale=0.3]{3peaks.png}
\caption{Three-Peak profiles for the off-center positions. 
   \label{fig:3p_profiles}} 
\end{figure*}

@@

\begin{figure*}
\includegraphics[scale=0.3]{3p_res.png}
\caption{Three-Peak profiles for different resolutions. 
   \label{fig:3p_res}} 
\end{figure*}

>>>>

A quantitative description of the \ly line has been reached through
Monte Carlo simulations \citep{Auer68,Avery68,Adams72}. In the last
two decades these studies have become popular due to the
availability of computing power. Early into the 21st century, the first
studies focused on homogeneous and static media
\citep{Ahn00,Ahn01,Zheng02}; Later on, the effects of clumpy media
\citep{Hansen06} and expanding/contracting shell/spherical geometries started to
be studied \citep{Verhamme06,Dijkstra06}. Similar codes have applied
these results to semi-analytic models of galaxy formation \citep{Orsi12} and
results of large hydrodynamic simulations
\citep{CLARA,Forero12,Behrens13}. Recently, Monte Carlo codes have also
been applied to the results of high resolution hydrodynamic
simulations of individual
galaxies\citep{Laursen09,Barnes11,Verhamme12,Yajima12}. Meanwhile, recent
developments have been focused on the systematic study of clumpy
outflows \citep{DijkstraKramer}and anisotropic velocity configurations
\citep{Zheng2013}.

@@

Performing that study is the main goal of this paper. We investigate for the
first time the impact of rotation on the morphology of the \ly line. We focus on a simplified system: a spherical gas cloud
with homogeneous density and solid body rotation, to study the line
morphology and the escape fraction in the presence of dust. We base
our work on two independent Monte Carlo based radiative transfer codes
CLARA \citep{CLARA} and XX \citep{DijkstraKramer} .

This paper is structured as follows: In \S
\ref{sec:implementation} we present the implementation of bulk
rotation into the Monte Carlo codes, paying special attention to coordinate
definitions. We also list the different physical parameters in
the simulated grid of models. In the next \S \ref{sec:results} we present the
results of the simulations, with special detail to quantities in the
line that show a clear evolution as a function of the sphere
rotational velocity. In \S \ref{sec:discussion} we discuss the
implications of our results in the interpretation of LAEs
observations and further refinements that can be implemented in the
theoretical modeling of the \ly line. In the last section we present
our conclusions.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NewWidthVvsVmax.png}
    \caption{Half-width for the non-dusty models as a function of
      rotational velocity $V_{\rm max}$. Continuous (dashed) lines
      correspond to central (homogeneous) source
      distributions. \label{fig:widthvsvelocity}} 
\end{figure}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{maximumVvsVmax.png}
\caption{Position of the peak maxima as a function of rotational
  velocity $V_{\rm max}$. Continuous (dashed) lines correspond to
  central (homogeneous) source distributions. A value of $x_{\rm
    max}=0$ indicates that line becomes single
  peaked. \label{fig:maximumsvsvelocity}}  
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Logarithm of the average number of scatterings as function of
  rotational velocity velocity. Continuous (dashed) lines represent an
  central (homogeneous) distribution of sources. \label{fig:Nscatt}}    
\end{figure}

@@

\begin{figure}
  \includegraphics[width=0.45\textwidth]{escapefraction.png}
   \caption{Escape fraction as a function of rotational velocity, for these models we took $\tau_{a}=1$. The
     continuous (dashed) lines correspond to central (homogeneous)
     models.
     \label{fig:efvsv}}
\end{figure}

@@

\begin{figure}
  \includegraphics[scale=0.35]{muhisto.png}
  \caption{$\mu$ histogram showing the orientation effects for the different
  distributions central(Points line), homogeneous(Dashed) and off-Center(Solid line).  
    \label{fig:muhisto}} 
\end{figure}

@@

\begin{figure*}
\includegraphics[scale=0.3]{OC_profiles.png}
\caption{\ly profiles for the off-center positions for all of our models, Dashed/solid
  lines represents spheres centered at $-0,5R/+0.5R$ respectively.
   \label{fig:OC_profiles}} 
\end{figure*}

@@

\begin{figure*}
\includegraphics[scale=0.3]{3peaks.png}
\caption{Three-Peak profiles for three different off-center positions $C=0.25R$ (doted line), $C=0.5R$ (dashed line) \& $C=0.75R$ (solid line) and for different rotational velocities. 
   \label{fig:3p_profiles}} 
\end{figure*}

@@

\begin{figure*}
\includegraphics[scale=0.3]{3p_res.png}
\caption{Three-Peak profiles for different resolutions:Low (Solid), medium(dashed), high(doted), 
   \label{fig:3p_res}} 
\end{figure*}
